### 2013-08-04 man/nips-2013-rebuttal.md
First draft of rebuttal

# NIPS 2013 Rebuttal

**Paper ID**: 1352  
**Title**: Mental Rotation as Bayesian Quadrature

## Reviewer 3

Reviewer 3 was primarily concerned with the comparison between models
and human data. While we did not collect human data ourselves, we
compared our model to the human response times reported by Shepard and
Metzler [6] (the key comparison is between Figure 1, depicting the
classic results, and Figure 5, illustrating our modeling results).
The mental rotation results are some of the most famous findings in
cognitive science ([6] alone has at least 2,061 citations according to
Web of Knowledge; 3,640 according to Google Scholar) and have been
replicated many times with many types of stimuli (Thomas 2013). As a
proof-of-concept, we demonstrate that our model is qualitatively
similar to the standard mental rotation result.  No other
computational models have been proposed which can explain how people
choose the correct direction and duration of rotation.

The reviewer was also concerned about how the prior over stimuli
related to the process by which the shapes were created. To clarify,
the shapes were generated in a manner consistent with the prior:
vertex angles were chosen uniformly at random between 0 and $2\pi$,
and vertex radii were chosen uniformly at random between 0 and 1.
Edges were drawn between vertices which were adjacent when sorted by
angle.

## Reviewer 4

Reviewer 4 was concerned with the explanatory power of our model,
suggesting that people do not need to integrate over the full set of
rotations. However, it is not obvious how the strategy "just rotate
until they find a match that is easy to recognize or exhaust the set
of potential rotations" should be implemented. Similarly, stopping
when a local maximum is found is not equivalent to quitting "even if
it's obviously not a match at that rotation". What does it mean for a
match to be easy to recognize, or for something be obviously not a
match? A perfect match cannot truly be identified due to perceptual
uncertainty in people's visual system. Moreover, one cannot just
rotate until a threshold of similarity has been reached, because the
similarity function is not normalized and therefore the value of its
global maximum is not known *a priori*.

If the value of the global maximum of the similarity function is not
known, then it is impossible to know if any particular maximum is the
global maximum or not. To reliably find the global maximum, then,
would require an exhaustive search (a full rotation of 360 degrees),
which is inconsistent with the linear nature of human response times
from [6]. Thus, it is necessary to either be satisfied with a local
maximum (as found by hill-climbing), or by a more sophisticated method.

Even if it was possible to perfectly recognize a match after rotating
the appropriate amount, there is still a separate issue: how do people
know which direction to rotate? Simply choosing one direction at
random and rotating to the true angle produces a pattern of data which
is inconsistent with the behavioral data from [6]. For any true
minimum angle $r$, we would have expected response time proportional
to $E[r] = 0.5r + 0.5(2\pi - r) = \pi$, which is constant rather than
monotonic.

The "model rotations" described around line 372 and shown in Figure 5
do refer to the number of steps the models takes. A more precise label
would be "total distance rotated, in degrees" which is directly
proportional to the number of steps the model takes.

To address the reviewer's question about whether the behavioral
results hold for our stimuli, we emphasize (as above in response to
Reviewer 3) that the Shepard and Metzler results [6] are very robust:
they have been replicated many times and with many other types of
stimuli. In particular, our stimuli are similar to those used by [23].

The reviewer correctly notes that the 95% confidence threshold used by
the BQ model's hypothesis test is likely the source of the 4% error
rate. The reviewer also points out that the MSE on line 382 is
inconsistent with our definition of MSE. This is indeed a mistake: the
MSE cannot be normalized between 0 and 1 because it has no upper
bound. The correct, non-normalized MSEs are 0.03 for the naive model,
0.002 for the parametric model, and 0.001 for the nonparametric model.

## Reviewer 5

We thank Reviewer 5 for their encouraging comments. We do intend to
collect human data in the future and to perform a more careful,
quantitative comparison between the model and participants. We further
hope that the model will be useful in uncovering previously unexamined
nuances in people's behavior.

Regarding people's mental representations, we do not think that people
are necessarily performing full Bayesian quadrature in their
heads. Ours is a computational-level model [3], which means that it is
a formal way of describing how the problem of determining if two
objects are the same could be solved. Although we incorporate the
constraint of sequential rotations, this is not a true
algorithmic-level constraint, as we have not given specific limits on
working memory, processing power, knowledge representation, etc. The
algorithmic-level form of this solution would be an approximation to
our computational-level model and would make the resource limitations
explicit. Currently, however, we have not explored possible
algorithmic approximations. Doing so is a direction for future work.

# References

* Thomas, Nigel J.T., "Supplement: Mental Rotation", *The Stanford
  Encyclopedia of Philosophy*, Edward N. Zalta (ed.). Retrieved on
  08/05/2013 from
  http://plato.stanford.edu/archives/spr2013/entries/mental-imagery/mental-rotation.html

* [6] R. N. Shepard and J. Metzler, "Mental Rotation of
  Three-Dimensional Objects," *Science*, vol. 171, no. 3972,
  pp. 701–703, 1971.

* [23] L. A. Cooper, "Mental Rotation of Random Two-Dimensional
  Shapes," *Cognitive Psychology*, vol. 7, pp. 20-43, 1975.