### 2013-10-09 man/nips-2013-bayesian-optimization.tex
Begin work on bayesian optimization submission

\documentclass{article} % For LaTeX2e
\usepackage{nips12submit_e,times}
%\documentstyle[nips12submit_09,times,art10]{article} % For LaTeX 2.09

\usepackage{amsmath, amsthm, amssymb}
% \usepackage{natbib}
\usepackage{graphicx}
\usepackage{wrapfig}
\usepackage{subcaption}

\pagestyle{empty}

\title{Mental Rotation as Bayesian Quadrature}
\author{Jessica B.~Hamrick and Thomas L.~Griffiths\\
  Department of Psychology\\
  University of California, Berkeley\\
  Berkeley, CA 94720\\
  \texttt{\{jhamrick,tom\_griffiths\}@berkeley.edu}}

% The \author macro works with any number of authors. There are two commands
% used to separate the names and addresses of multiple authors: \And and \AND.
%
% Using \And between authors leaves it to \LaTeX{} to determine where to break
% the lines. Using \AND forces a linebreak at that point. So, if \LaTeX{}
% puts 3 of 4 authors names on the first line, and the last on the second
% line, try using \AND instead of \And before the third author name.

\newcommand{\TODO}[1]{\textcolor{red}{[TODO: #1]}}
\newcommand{\ME}[0]{\mathrm{ME}}
\newcommand{\naive}[0]{na\"ive}
\newcommand{\Naive}[0]{Na\"ive}

\include{nips-2013-analyses}

\nipsfinalcopy % Uncomment for camera-ready version

\begin{document}

\maketitle

\begin{abstract}
  Given a computational resource--for example, the ability to
  visualize an object rotating--how do you best make use of it? We
  explored how mental simulation should used in the classic
  psychological task of determining if two images depict the same
  object in different orientations. We compared three models on this
  mental rotation task, and found that a model based on an optimal
  experiment design for Bayesian quadrature is objectively more
  accurate and qualitatively more consistent with classic behavioral
  data than the other models we evaluated. We suggest that rational
  models which adaptively exploit available resources are promising in
  their ability to characterize metacognitive processes like mental
  simulation.
\end{abstract}

\section{Introduction}

One of the challenges of solving any computational problem is
determining how best to use the available computing resources. For
example, a computer can render complex graphics faster by recognizing
that this kind of computation should be carried out by a specialized
graphics processor. The same challenge arises in designing an
intelligent agent: how should the agent make best use of its computing
resources? Recent research on rational models of human cognition has
provided insight into the nature of the computational problems that
human beings need to solve (e.g., \cite{Chater:1999wp,tenenbaumkgg11}),
but leaves open the question of how people allocate their resources in
solving those problems. In this paper, we take a step towards
addressing this question, applying rational analysis (in the spirit of
\cite{Marr:1983to,anderson90,Shepard:1987tt}) to one aspect of human
metacognition: the use of mental simulation.

Consider the images on the left in Figure
\ref{fig:mental-rotation}. In each panel, are the two depicted objects
identical (except for a rotation), or distinct? When presented with
this mental rotation task, people default to a strategy in which they
visualize one object rotating until it is congruent with the other
\cite{Shepard1971}. There is strong evidence for such mental
simulation: we can imagine three-dimensional objects in our minds and
manipulate them, to a certain extent, as if they were real
\cite{Kosslyn:2009tj}.  However, the use of mental simulation is
predicated on determining appropriate parameters to give the
simulation, analogous to determining exactly what computation should
be passed to a graphics processor.  In the case of the classic mental
rotation task, we might ask: How do people know which way to rotate
the object?  When should one stop rotating and accept the hypothesis
that the objects are different?

Recent work in cognitive science has shown how problems of allocating
cognitive resources to solving computational problems can be analyzed
using the methods of statistical decision theory
\cite{Lieder:2012wg,Vul:2009wy}. We apply this ``rational
metacognition'' approach to the problem of mental rotation. We
investigate several computational solutions to this task and
qualitatively compare them to human mental rotation performance. In
particular, we argue that performing mental rotation can be framed as
integration over a probability distribution, with the direction of
rotation becoming an optimal experiment design problem (or in machine
learning parlance, an active learning problem). We show that recent
work on methods for Bayesian quadrature
\cite{Diaconis:1988uo,OHagan:1991tx,Osborne:2012tm} provides a way to
solve this problem, outperforming simpler heuristics for determining
the direction and extent of rotation.

The plan of the paper is as follows. First, we present the problem
domain and a computational-level analysis of how to solve it. We
describe two different models which approximate this solution,
evaluate their accuracy at approximating the computational solution,
and qualitatively compare their behavior to that of the classic mental
rotation results. We end with a brief summary and conclusion of our
approach.

\begin{figure}[t]
  \centering
  \includegraphics[width=0.7\textwidth]{../figures/shepard-rotation.png}
  \caption{\textbf{Classic mental rotation task}. Participants in
    \cite{Shepard1971} saw stimuli such as those on the left, and
    judged whether each pair of shapes was the same shape in two
    different orientations (``same'' pairs), or two different shapes
    (``different'' pairs). \textbf{A} and \textbf{B} show ``same''
    pairs, while \textbf{C} shows a ``different'' pair. The plots on
    the right indicate mean response times on the mental rotation
    task, exhibiting a strong linear relationship with increasing
    variance as a function of true rotation.}
  \label{fig:mental-rotation}
\end{figure}

% \section{Background}

% There is a long and rich history of work on mental rotation beginning
% with Shepard and Metzler \cite{Shepard1971}, who presented
% participants with pairs of images such as those shown in Figure
% \ref{fig:mental-rotation} (left panel). These pairs of
% three-dimensional objects could either be: the same object rotated in
% the plane (Figure \ref{fig:mental-rotation}A), the same object rotated
% in depth (Figure \ref{fig:mental-rotation}B), or two different objects
% (Figure \ref{fig:mental-rotation}C; in this case, the ``different''
% objects had reflective symmetry, but no rotational
% symmetry). Participants had to determine whether the two images were
% of the same object or not, and Shepard and Metzler famously found that
% response times for plane and depth rotations had a strong linear
% relationship with the minimum angle of rotation between the two
% objects (Figure \ref{fig:mental-rotation}, right panel). The
% conclusion was that participants were visually ``rotating'' the
% objects in their minds.

% For many years, this idea was contested, with some researchers arguing
% the underlying cognitive processes were not visual in nature (e.g.,
% \cite{Pylyshyn1981}). In particular, Anderson \cite{Anderson1978}
% proved that the mental imagery debate could not be resolved on the
% basis of response time data alone. In recent years, however,
% significant efforts have investigated the neural underpinnings of the
% mental rotation phenomena, arguing that the results of brain imaging
% studies support the existence of visual imagery
% \cite{Kosslyn:2009tj,Kosslyn1988}.  We assume that mental imagery is
% indeed a visual process, and turn to the question of how, and when, it
% is used. Previous work has examined how people might use imagery
% \cite{Hegarty2004, Schwartz1999} to solve reasoning problems, and when
% they might use imagery as opposed to a symbolic rule
% \cite{Schwartz:1996uy}. Eye-tracking studies of people performing
% mental rotations suggest strategies that people might be using, such
% as aligning smaller features of the images \cite{Just1976}. People
% seem to use imagery methodically, which raises the question: what is
% the method?

% Others studying people's pattern of behavior when engaged in
% self-directed learning have found that they tend to use a strategy of
% \textit{active learning} \cite{Gureckis:2012gu,Markant:2012uu}. When
% given a choice as to the information they can obtain, people will not
% choose the information randomly (as is the case with a \naive{} Monte
% Carlo sampler), but according to some utility function such as
% information gain \cite{Markant:2012uu,Nelson2007}. We propose that
% such a strategy can be combined with the notion of mental imagery as a
% tool: an optimal model of mental rotation may, perhaps, rely on an
% active learning strategy. In the next sections, we investigate this
% idea with several different approaches.

\section{Computational-level model}

We begin by first analyzing mental rotation at Marr's
\textit{computational} level \cite{Marr:1983to}: what is the problem
to be solved, and what is the optimal way to do so?  Formally, people
are presented with two images, $X_a$ and $X_b$, which are the
coordinates of the vertices of two-dimensional shapes similar to those
used by \cite{Cooper:1975wp} (e.g., Figure
\ref{fig:stimuli}). Participants must determine whether $X_a$ and
$X_b$ were generated from the same (albeit possibly transformed and
permuted) original shape, i.e., whether $\exists R,M\textrm{ s.t. }
X_b=MRX_a$, where $M$ is a permutation matrix and $R$ is a rotation
matrix.

We can formulate the judgment of whether $X_a$ and $X_b$ have the same
origins by deciding about two hypotheses, $h_0$: $\forall M,R\ X_b\neq
MRX_a$ and $h_1$: $\exists M,R\textrm{ s.t. } X_b=MRX_a$.  To compare
the hypotheses, we need to compute the posterior for each: $p(h\
\vert\ X_a, X_b)\propto p(X_a, X_b\ \vert\ h)p(h)$.

Assuming the hypotheses are equally likely \textit{a priori}, the
prior term $p(h)$ will cancel out when comparing $h_0$ and $h_1$, thus
allowing us to focus on the likelihoods:
\begin{align}
  p(X_a,\ X_b\ \vert \ h_0)&=p(X_a)p(X_b), \label{eq:lh-h0}\\
  p(X_a,\ X_b\ \vert \ h_1)&=\int_R\int_M p(X_a) p(X_b\vert X_a,R,M) p(R) p(M)\ \mathrm{d}M\ \mathrm{d}R. \label{eq:lh-h1}
\end{align}
Under $h_0$, the likelihood is easy to compute because we assume that
$X_a$ and $X_b$ are independent (Equation \ref{eq:lh-h0}). When the
configurations are the same, the likelihood becomes more complicated
(Equation \ref{eq:lh-h1}). For a small number of vertices, we can
compute the integral over $M$ by enumerating every possible mapping
between $X_a$ and $X_b$. After doing so, we obtain:
\begin{equation} 
  p(X_a,\ X_b\ \vert \ h_1)=\int_R p(X_a) p(X_b\vert X_a,R) p(R)\ \mathrm{d}R.
\end{equation}

Once we have computed both likelihoods, we compute their ratio $\ell$
which is given by $\ell=\left(\int_R p(X_b\ \vert\ X_a, R)p(R)\
  \mathrm{d}R\right) /\ p(X_b)$. If $\ell<1$, then $h_0$ is the more
likely hypothesis. If $\ell>1$, then $h_1$ is the more likely
hypothesis.

\begin{figure}[t]
  \centering
  \begin{subfigure}[b]{0.32\textwidth}
    \centering
    \includegraphics[width=\textwidth]{../figures/stimuli_shapes.pdf}
    \vspace{0pt}
    \caption{Example stimuli}
    \label{fig:stimuli}
  \end{subfigure}
  \begin{subfigure}[b]{0.32\textwidth}
    \centering
    \includegraphics[width=\textwidth]{../figures/li_regression.pdf}
    \caption{\Naive{} model}
    \label{fig:li}
  \end{subfigure}
  \begin{subfigure}[b]{0.32\textwidth}
    \centering
    \includegraphics[width=\textwidth]{../figures/bq_regression_final.pdf}
    \caption{Bayesian Quadrature Model}
    \vspace{0pt}
    \label{fig:bq}
  \end{subfigure}
  \caption{\textbf{(a)} An example stimulus in which the shapes differ
    only by a rotation. All stimuli consist of three to six vertices
    centered around the origin, and edges which create a closed loop
    from the vertices. The true angle of rotation between $X_a$ and
    $X_b$ is at $\frac{2\pi}{3}$. \textbf{(b-c)} Likelihood function
    and \naive{} (b) and BQ (c) model estimates. The sampled points
    $\mathbf{R}$ (red circles) are then used to estimate $S$ (black
    lines are the true $S$, red lines are the estimate).}
  \label{fig:shapes}
\end{figure}

\section{Metacognitive algorithms}

Given the computational model defined above, we now explore the ways
in which it can be expressed algorithmically. First, we define the
prior probabilities over stimuli according to a generative
procedure. A set of $n$ vertices could be chosen in any of $n!$
different ways, and each vertex is located at a random angle (between
0 and $2\pi$) and radius (between 0 and 1). Thus, the prior over a
shape $X$ is $p(X)=n!\left(\frac{1}{2\pi}\right)^n$, which gives us
the denominator of $\ell$. Computing the numerator is more difficult,
as we we cannot compute $p(X_b\vert X_a, R)$ directly. Instead, we
introduce a new variable $X_R$ denoting a mental image, which
approximates $RX_a$. The $X_R$ are generated by repeated application
of a function $\tau$:
\begin{equation}
  X_R=RX_a=\tau(X_{R-r}, r)=\tau(\tau(X_{R-2r}, r), r)=\ldots{}=\tau^{(\frac{R}{r})}(X_a, r).
  \label{eq:tau}
\end{equation} 
where $r$ is a small angle, and $\tau^{(i)}$ indicates $i$ recursive
applications of $\tau$. Using this sequential function, we get:
\begin{align}
  p(X_a, X_b\ \vert \ h_1)&=\int_R \int_{X} p(X_b\vert X) p(X\vert X_a, R)p(X_a)p(R)\ \mathrm{d}X\ \mathrm{d}R \nonumber \\
  &= \int_R \int_X p(X_b\vert X)\delta(\tau^{(\frac{R}{r})}(X_a, r)-X)p(X_a)p(R)\ \mathrm{d}X\ \mathrm{d}R \nonumber \\
  &= \int_R p(X_b\vert X_R)p(X_a)p(R)\ \mathrm{d}R
\end{align}
However, the exact form of $p(X_b\vert X_R)$ is still unknown. We
approximate it with a similarity function $S(X_b, X_R)$, and denote
the resulting integral as $Z$:
\begin{equation}
Z=\int_R S(X_b, X_R)p(R)\ \mathrm{d}R\approx \int_R p(X_b\vert X_R)p(R)\ \mathrm{d}R.
\label{eq:Z}
\end{equation}
We define the similarity function $S$, which incorporates the
different mappings as vertices, as follows. Because the vertices are
connected in a way which forms a closed loop, we need only consider
$2n$ mappings of the $n$ vertices (we assume uncertainty for which is
the ``first'' vertex, and then which of its two neighbors is the
``second''). So, the possible orderings are of the form $M=\lbrace{}0,
1, \ldots{}, n\rbrace{}$, $M=\lbrace{}n, 0, \ldots{}, n-1\rbrace{}$,
and so on. Combining this with a Gaussian similarity metric, we
obtain:
\begin{equation}
  S(X_b, X_R)=\frac{1}{2n}\sum_{M\in\mathbb{M}}\prod_{i=1}^n\mathcal{N}(X_b[i]\ \vert \ (MX_R)[i], \Sigma)
  \label{eq:similarity}
\end{equation}
where $i$ denotes the $i^{th}$ vertex. An example stimulus and
corresponding $S$ is shown in Figure \ref{fig:shapes}.

To summarize, the process of generating a mental image consists of
computing a single $X_R$ (as in Equation \ref{eq:tau}) and then
computing $S(X_b, X_R)$. We denote the sequence of rotations computed
by this procedure as $\mathbf{R}=\{R_1, R_2, \ldots{}\}$. However,
this sequence cannot be arbitrary, as mental rotation is
computationally demanding. Our goal is to minimize the number of
rotations $\vert\mathbf{R}\vert$ while still obtaining an estimate of
$Z$ that is accurate enough to choose the correct hypothesis. With
this in mind, we now examine several approaches to solve this problem.

\paragraph{\Naive{}}

As a baseline, we defined a \naive{} model which performs a
hill-climbing search over the similarity function until it reaches a
(possibly local) maximum. Once a maximum as been found, the model
computes an estimate of $Z$ by linearly interpolating between sampled
rotations. Figure \ref{fig:li} shows an example of the \naive{}
model's estimate of $S$.

\paragraph{Bayesian Quadrature}

A more flexible strategy uses what is known as \emph{Bayesian
  Quadrature} \cite{Diaconis:1988uo,OHagan:1991tx} to estimate $Z$.
Bayesian Quadrature (BQ) allows us to compute a posterior distribution
over $Z$ by placing a Gaussian Process (GP) prior on the function $S$
and evaluating $S$ at a particular set of points. Because our data is
circular, we use a periodic kernel \cite{Rasmussen:2006vz}.

\begin{figure}[t]
  \centering
  \includegraphics[width=\textwidth]{../figures/model_rotations.pdf}
  \caption{\textbf{Model rotations.} Top: each subplot shows the
    correspondence between the true angle of rotation ($R$) for
    ``same'' pairs and the amount of rotation performed by the
    model. Bottom: each subplot shows the models' mean rotations over
    stimuli pairs that were judged correctly. Black dots correspond to
    ``same'' pairs, and blue lines correspond to ``different'' pairs
    (for which the mean rotation is constant across true rotations, as
    the true rotation is undefined). Error bars/shaded regions
    indicate one standard deviation, and the dotted lines indicate the
    least-squares fit to the ``same'' pairs.}
  \label{fig:rotations}
\end{figure}

BQ has its difficulties, however. While in our case $S$ is a
non-negative likelihood function, GP regression enforces no such
constraint. In an effort to avoid this problem, \cite{Osborne:2012tm}
give a method to place a prior over the log likelihood\footnote{In
  practice, as in \cite{Osborne:2012tm}, we use the transform of
  $\log(S+1)$. Additionally, we follow \cite{Osborne:2012tm} and use a
  combination of MLII and marginalization to fit the $w$ kernel
  parameters. The output scales are fixed at $h_S=\sqrt{0.1}$ and
  $h_{\log S}=\log(h_S + 1)$.}, thus ensuring that $S=e^{\log S}$ will
be positive:
\begin{equation*}
  E[Z\ \vert \ \log S]=\int_{\log S}\left(\int_R \exp(\log{S(X_b,X_R)})p(R)\ \mathrm{d}R\right)\mathcal{N}\left(\log{S}\ \vert \ \mu_{\log S}, \Sigma_{\log S}\right)\ \mathrm{d}\log S
\end{equation*}
where $\mu_{\log S}$ and $\Sigma_{\log S}$ are the mean and
covariance, respectively, of the GP regression over $\log S$ given
$\mathbf{R}$. We approximate this according to the method given in
\cite{Osborne:2012tm}:
\begin{equation}
  \mu_Z=E[Z\ \vert \ S, \log S, \Delta_c] \approx \int_R \mu_{S}(1 + \mu_{\Delta_c}) p(R)\ \mathrm{d}R 
  \label{eq:bq-Z-mean}
\end{equation}
where $\mu_S$ is the mean of a GP regression over $S$ given
$\mathbf{R}$; and $\mu_{\Delta_c}$ is a regression over
$\Delta_c=\mu_{\log S} - \log \mu_S$ given $\mathbf{R}_c$, which
consists of $\mathbf{R}$ and a set of intermediate \emph{candidate
  points} $c$ as described in \cite{Osborne:2012tm}. The variance is
$\tilde{V}(Z\vert S, \log S, \Delta_c)$ as defined in Equation 12 of
\cite{Osborne:2012tm}.

To start, we pick an initial direction of rotation which results in
the higher value of $S$. At each step, we compute $\mu_Z$ and
$\tilde{V}$ in order to estimate a distribution over the likelihood
ratio $\ell$, i.e.  $p(\ell)\approx\frac{1}{p(X_b)}\ \mathcal{N}(Z\
\vert\ \mu_Z, \sigma_z)$.  We choose $h_0$ when $p(\ell < 1)\geq
0.95$, and $h_1$ when $p(\ell > 1)\geq 0.95$. Until one of these
conditions are met (or the shape has been fully rotated), the model
will continues to compute rotations and update its estimate of $Z$.

We additionally allow the model to change direction or ``reset'' based
on an estimate of the posterior variance of $Z$ given some new sample
$a$. This is similar to the procedure given in \cite{Osborne:2012tm},
however we do not compute the full posterior variance. Instead, we
compute the variance given only the current mean estimate of $a$. If
this estimated variance is lowered more by ``resetting'', then the
model will change directions. Thus, it is able to have an active say
in which mental rotations are computed, unlike the hill-climbing
procedure.

\section{Results}

We evaluated each model's performance on 20 shapes which had between 3
and 6 vertices, inclusive, (e.g., $X_a$ in Figure
\ref{fig:stimuli}). For each shape, we computed 18 ``same'' and 18
``different'' stimuli pairs, with $R$ spaced at $20^\prime$ increments
between 0 and 360, as in \cite{Shepard1971}. ``Same'' pairs were
created by rotating $X_a$ by $R$; the same was true for ``different'',
except that $X_a$ was also reflected across the $y$-axis. Each shape
was generated randomly, but under the constraint that the
``same''/``different'' pairs created from the shape were
distinguisable.

We considered two metrics of performance. First, we looked at
\textit{response error rates}: How accurate was the model at choosing
the correct hypothesis? This was defined as the mean error ($\ME{}$),
or fraction of times the model chose incorrectly.

We also looked at \textit{rotations}: For those ``same'' pairs which
the model judged correct, how correlated were the model's rotations
with the true angles of rotation?  We quantified this using the
Pearson's correlation coefficient $\rho$ for the true rotation ($R$)
versus the extent of the model's rotations ($\vert
\mathbf{R}\vert$). The top row in Figure \ref{fig:rotations} shows
individual points corresponding to true rotations vs. rotations by the
models for all stimuli, including those judged incorrectly. The bottom
row shows mean model rotations, across only those stimuli which were
judged correctly. We structured the analysis in this way to better
compare to the results of \cite{Shepard1971}, which also excluded
incorrectly judged stimuli.

\paragraph{\Naive{}}

The \naive{} model's response error rate was $\NaiveME{}$, which is
better than chance (equivalent to guessing randomly,
i.e. $\ME{}=0.5$). The correlation between the \naive{} model's
average rotation and the true angle of rotation was $\Naivecorr{}$
(Figure \ref{fig:rotations}, bottom left). More complex patterns are
revealed by examining the shape of the raw data in Figure
\ref{fig:rotations}, top left: the \naive{} model corresponds
extremely well to the true angle of rotation for approximately
$R<\frac{\pi}{2}$. This is unsurprising, because the closer the true
angle is to zero, the less the model has to rotate, and the less
likely it will get stuck on local maxima. Thus, it is more likely to
locate the global maximum, which corresponds to the true angle of
rotation. For $R>\frac{\pi}{2}$, we see an increasing tendency to
under-rotate due to getting stuck on local maxima, as well as a
tendency to over-rotate if the wrong direction was initially chosen.

\paragraph{Bayesian Quadrature}

The BQ model was much more accurate in choosing the correct hypothesis
than the other two models ($\BQME{}$). The average rotations computed
by the BQ model were strongly correlated with the true rotations
($\BQcorr$, see Figure \ref{fig:rotations}, bottom right). Because the
BQ model has the capacity to ``reset'', it could recover from rotating
in the incorrect direction (e.g., Figure \ref{fig:bq}) and thus did
not over-rotate as frequently.  It also under-rotated less frequently:
because the model continues to rotate until it is confident that its
estimate of $Z$ is accurate, it does not get stuck as easily on local
optima.

\section{Conclusion}

In this paper, we asked: how do people use mental simulation?  We
investigated the specific case of mental rotation, using rational
analysis to characterize optimal strategies for performing mental
rotation. We found that an adaptive, BQ model performs best at the
classic rotation task \cite{Shepard1971}, as it is able to actively
monitor the confidence of its estimate and intelligently choose the
direction of rotation. Two simpler models based on heuristics
performed much worse: they did not maintain the linear relationship
with the true angle of rotation due to lack of robustness against
local maxima or incorrect rotation direction.  More broadly, the BQ
model provides answers to puzzling questions surrounding the
incremental nature of mental rotation: which way should the object be
rotated, and for how long? This model formalizes these answers in a
way that is qualitatively consistent with human behavior, both in
response time linearity \cite{Shepard1971} and variability
\cite{Just1976}. These results further indicate that this may be
another instance in which people do appropriately use available
computational resources to solve the task at hand.  Rational, adaptive
approaches such as this may therefore be successful when applied to
more complex cognitive processes above and beyond mental rotation.

% two questions: which way to rotate, when to stop, giving formal answer
% to these questions which produces answers consistent with human
% behavior. instance of showing people are appropriately using
% computationl resources

% \textbf{Acknowledgments}
% \TODO{}

% \renewcommand\bibsection{\subsubsection*{\refname}}
\renewcommand\refname{\normalsize{References}}
\bibliographystyle{ieeetr}
{\small \bibliography{references}}

\end{document}