### 2013-10-15 man/nips-2013-bayesian-optimization.tex
Trim down the submission to 4 pages plus references

\documentclass{article} % For LaTeX2e
\usepackage{nips12submit_e,times}
%\documentstyle[nips12submit_09,times,art10]{article} % For LaTeX 2.09

\usepackage{amsmath, amsthm, amssymb}
% \usepackage{natbib}
\usepackage{graphicx}
\usepackage{wrapfig}
\usepackage{subcaption}

\pagestyle{empty}

\title{Mental Rotation as Bayesian Quadrature}
\author{Jessica B.~Hamrick and Thomas L.~Griffiths\\
  Department of Psychology, University of California, Berkeley\\
  Berkeley, CA 94720\\
  \texttt{\{jhamrick,tom\_griffiths\}@berkeley.edu}}

% The \author macro works with any number of authors. There are two commands
% used to separate the names and addresses of multiple authors: \And and \AND.
%
% Using \And between authors leaves it to \LaTeX{} to determine where to break
% the lines. Using \AND forces a linebreak at that point. So, if \LaTeX{}
% puts 3 of 4 authors names on the first line, and the last on the second
% line, try using \AND instead of \And before the third author name.

\newcommand{\TODO}[1]{\textcolor{red}{[TODO: #1]}}
\newcommand{\ME}[0]{\mathrm{ME}}
\newcommand{\naive}[0]{na\"ive}
\newcommand{\Naive}[0]{Na\"ive}

\include{nips-2013-analyses}

\nipsfinalcopy % Uncomment for camera-ready version

\begin{document}

\maketitle

\begin{abstract}
  Given a computational resource--for example, the ability to
  visualize an object rotating--how do you best make use of it? We
  explored how mental simulation should used in the classic
  psychological task of determining if two images depict the same
  object in different orientations. We compared two models on this
  mental rotation task, and found that a model based on an optimal
  experiment design for Bayesian quadrature is objectively more
  accurate and qualitatively more consistent with classic behavioral
  data than a simpler model. We suggest that rational models which
  adaptively exploit available resources are promising in their
  ability to characterize metacognitive processes like mental
  simulation.
\end{abstract}

\section{Introduction}

One of the challenges of solving any computational problem is
determining how best to use the available computing resources. For
example, a computer can render complex graphics faster by recognizing
that this kind of computation should be carried out by a specialized
graphics processor. The same challenge arises in designing an
intelligent agent: how should the agent make best use of its computing
resources? Recent research on rational models of human cognition has
provided insight into the nature of the computational problems that
human beings need to solve (e.g., \cite{Chater:1999wp,tenenbaumkgg11}),
but leaves open the question of how people allocate their resources in
solving those problems. In this paper, we take a step towards
addressing this question, applying rational analysis (in the spirit of
\cite{Marr:1983to,anderson90,Shepard:1987tt}) to one aspect of human
metacognition: the use of mental simulation.

Consider the images on the left in Fig.
\ref{fig:mental-rotation}. In each panel, are the two depicted objects
identical (except for a rotation), or distinct? When presented with
this mental rotation task, people default to a strategy in which they
visualize one object rotating until it is congruent with the other
\cite{Shepard1971}. There is strong evidence for such mental
simulation: we can imagine three-dimensional objects in our minds and
manipulate them, to a certain extent, as if they were real
\cite{Kosslyn:2009tj}.  However, the use of mental simulation is
predicated on determining appropriate parameters to give the
simulation, analogous to determining exactly what computation should
be passed to a graphics processor.  In the case of the classic mental
rotation task, we might ask: How do people know which way to rotate
the object?  When should one stop rotating and accept the hypothesis
that the objects are different?

Recent work in cognitive science has shown how problems of allocating
cognitive resources to solving computational problems can be analyzed
using the methods of statistical decision theory
\cite{Lieder:2012wg,Vul:2009wy}. We apply this ``rational
metacognition'' approach to the problem of mental rotation. We
investigate several computational solutions to this task and
qualitatively compare them to human mental rotation performance. In
particular, we argue that performing mental rotation can be framed as
integration over a probability distribution, with the direction of
rotation becoming an optimal experiment design problem (or in machine
learning parlance, an active learning problem). We show that recent
work on methods for Bayesian quadrature
\cite{Diaconis:1988uo,OHagan:1991tx,Osborne:2012tm} provides a way to
solve this problem, outperforming simpler heuristics for determining
the direction and extent of rotation.

\begin{figure}[t]
  \centering
  \includegraphics[width=0.9\textwidth]{../figures/shepard-rotation.png}
  \caption{\textbf{Classic mental rotation task}. Participants in
    \cite{Shepard1971} saw stimuli such as those on the left, and
    judged whether each pair of shapes was the same shape in two
    different orientations (``same'' pairs), or two different shapes
    (``different'' pairs). \textbf{A} and \textbf{B} show ``same''
    pairs, while \textbf{C} shows a ``different'' pair. The plots on
    the right indicate mean response times on the mental rotation
    task, exhibiting a strong linear relationship with increasing
    variance as a function of true rotation.}
  \label{fig:mental-rotation}
\end{figure}

\section{Computational-level model}

We begin by first analyzing mental rotation at Marr's
\textit{computational} level \cite{Marr:1983to}: what is the problem
to be solved?  Formally, people are presented with two images, $X_a$
and $X_b$, which are the coordinates of the vertices of 2D shapes
(e.g., Fig.  \ref{fig:stimuli}). Participants must determine whether
$X_a$ and $X_b$ depict the same shape, i.e., whether $\exists
R\textrm{ s.t. } X_b=RX_a$, where $R$ is a rotation matrix. We can
formulate the judgment of whether $X_a$ and $X_b$ have the same
origins by deciding about two hypotheses, $h_0$: $\forall R\ X_b\neq
RX_a$ and $h_1$: $\exists R\textrm{ s.t. } X_b=RX_a$.  To compare the
hypotheses, we need to compute the posterior for each: $p(h\ \vert\
X_a, X_b)\propto p(X_a, X_b\ \vert\ h)p(h)$. Assuming the hypotheses
are equally likely \textit{a priori}, the prior term $p(h)$ will
cancel out when comparing $h_0$ and $h_1$, thus allowing us to focus
on the likelihoods, which are $p(X_a,\ X_b\ \vert \ h_0)=p(X_a)p(X_b)$
for $h_0$ and $p(X_a,\ X_b\ \vert \ h_1)=\int_R p(X_a) p(X_b\vert
X_a,R) p(R)\ \mathrm{d}R$ for $h_1$. From these likelihoods, we
compute the ratio $\ell$ which is given by $\ell=\left(\int_R p(X_b\
  \vert\ X_a, R)p(R)\ \mathrm{d}R\right) /\ p(X_b)$. If $\ell<1$, then
$h_0$ is the more likely hypothesis. If $\ell>1$, then $h_1$ is the
more likely hypothesis.

\section{Algorithmic approximation}

We define the prior probability of shape $X$ to be
$p(X)=n!\left(\frac{1}{2\pi}\right)^n$ according to a generative
procedure\footnote{A set of $n$ vertices could be chosen in any of
  $n!$ different ways, and each vertex is located at a random angle
  (between 0 and $2\pi$) and radius (between 0 and 1).}. This gives us
the denominator of $\ell$. Computing the numerator is more difficult,
as we we cannot compute $p(X_b\vert X_a, R)$ directly. Instead, we
introduce a new variable $X_R$ denoting a mental image, which
approximates $RX_a$. The $X_R$ are generated by repeated application
of a function $\tau$, i.e. $X_R=RX_a=\tau(X_{R-r},
r)=\tau(\tau(X_{R-2r}, r), r)=\ldots{}=\tau^{(\frac{R}{r})}(X_a, r)$,
where $r$ is a small angle, and $\tau^{(i)}$ indicates $i$ recursive
applications of $\tau$. Using this sequential function, we get:
\begin{align}
  &p(X_a, X_b\ \vert \ h_1)=\int_R \int_{X} p(X_b\vert X) p(X\vert X_a, R)p(X_a)p(R)\ \mathrm{d}X\ \mathrm{d}R \nonumber \\
  &= \int_R \int_X p(X_b\vert X)\delta(\tau^{(\frac{R}{r})}(X_a, r)-X)p(X_a)p(R)\ \mathrm{d}X\ \mathrm{d}R = \int_R p(X_b\vert X_R)p(X_a)p(R)\ \mathrm{d}R
\end{align}
However, the exact form of $p(X_b\vert X_R)$ is still unknown. We
approximate it with a similarity function $S(X_b, X_R)$, and denote
the resulting integral as $Z=\int_R S(X_b, X_R)p(R)\
\mathrm{d}R\approx \int_R p(X_b\vert X_R)p(R)\ \mathrm{d}R$.  We
define the similarity based on Gaussian similarity and possible
mappings $M$ of the vertices\footnote{Because the vertices are
  connected in a way which forms a closed loop, we need only consider
  $2n$ mappings of the $n$ vertices (we assume uncertainty for which
  is the ``first'' vertex, and then which of its two neighbors is the
  ``second''). So, the possible orderings are of the form
  $M=\lbrace{}0, 1, \ldots{}, n\rbrace{}$, $M=\lbrace{}n, 0, \ldots{},
  n-1\rbrace{}$, etc.}, i.e. $S(X_b,
X_R)=\frac{1}{2n}\sum_{M\in\mathbb{M}}\prod_{i=1}^n\mathcal{N}(X_b[i]\
\vert \ (MX_R)[i], \Sigma)$ where $i$ denotes the $i^{th}$ vertex. An
example stimulus and corresponding $S$ is shown in Fig.
\ref{fig:shapes}.

To summarize, the process of generating a mental image consists of
computing a single $X_R$ and then computing $S(X_b, X_R)$. We denote
the sequence of rotations computed by this procedure as
$\mathbf{R}=\{R_1, R_2, \ldots{}\}$. However, this sequence cannot be
arbitrary, as mental rotation is computationally demanding. Our goal
is to minimize the number of rotations $\vert\mathbf{R}\vert$ while
still obtaining an estimate of $Z$ that is accurate enough to choose
the correct hypothesis.

\paragraph{\Naive{}}

We defined a \naive{} model which performs a hill-climbing search over
the similarity function until it reaches a (possibly local)
maximum. Once a maximum as been found, the model computes an estimate
of $Z$ by linearly interpolating between sampled rotations (e.g.,
Fig. \ref{fig:li}).

\paragraph{Bayesian Quadrature}

A more flexible strategy uses what is known as \emph{Bayesian
  Quadrature} (BQ) \cite{Diaconis:1988uo,OHagan:1991tx} to estimate
$Z$.  BQ computes a posterior distribution over $Z$ by placing a
Gaussian Process (GP) prior on the function $S$ and evaluating $S$ at
a particular set of points. However, althought $S$ is a non-negative
likelihood function, GP regression enforces no such
constraint. \cite{Osborne:2012tm} give a method to place a prior over
the log likelihood\footnote{In practice, as in \cite{Osborne:2012tm},
  we use the transform of $\log(S+1)$. Additionally, we follow
  \cite{Osborne:2012tm} and use a combination of MLII and
  marginalization to fit the $w$ kernel parameters. The output scales
  are fixed at $h_S=\sqrt{0.1}$ and $h_{\log S}=\log(h_S + 1)$.}, thus
ensuring that $S=e^{\log S}$ will be positive, i.e.  $E[Z\ \vert \
\log S]=\int_{\log S}\left(\int_R \exp(\log{S(X_b,X_R)})p(R)\
  \mathrm{d}R\right)\mathcal{N}\left(\log{S}\ \vert \ \mu_{\log S},
  \Sigma_{\log S}\right)\ \mathrm{d}\log S$, where $\mu_{\log S}$ and
$\Sigma_{\log S}$ are the mean and covariance, respectively, of the GP
over $\log S$ given $\mathbf{R}$. We approximate this according to the
method given in \cite{Osborne:2012tm}, i.e. $\mu_Z=E[Z\ \vert \ S,
\log S, \Delta_c] \approx \int_R \mu_{S}(1 + \mu_{\Delta_c}) p(R)\
\mathrm{d}R$, where $\mu_S$ is the mean of a GP over $S$ given
$\mathbf{R}$; and $\mu_{\Delta_c}$ of a GP over $\Delta_c=\mu_{\log S}
- \log \mu_S$ given $\mathbf{R}_c$, which consists of $\mathbf{R}$ and
a set of intermediate \emph{candidate points} $c$ as described in
\cite{Osborne:2012tm}. The variance is $\tilde{V}(Z\vert S, \log S,
\Delta_c)$ as defined in Equation 12 of \cite{Osborne:2012tm}.

\begin{figure}[t]
  \centering
  \begin{subfigure}[b]{0.32\textwidth}
    \centering
    \includegraphics[width=\textwidth]{../figures/stimuli_shapes.pdf}
    \vspace{0pt}
    \caption{Example stimuli}
    \label{fig:stimuli}
  \end{subfigure}
  \begin{subfigure}[b]{0.32\textwidth}
    \centering
    \includegraphics[width=\textwidth]{../figures/li_regression.pdf}
    \caption{\Naive{} model}
    \label{fig:li}
  \end{subfigure}
  \begin{subfigure}[b]{0.32\textwidth}
    \centering
    \includegraphics[width=\textwidth]{../figures/bq_regression_final.pdf}
    \caption{Bayesian Quadrature Model}
    \vspace{0pt}
    \label{fig:bq}
  \end{subfigure}
  \caption{\textbf{(a)} An example stimulus in which the shapes differ
    only by a rotation. All stimuli consist of three to six vertices
    centered around the origin, and edges which create a closed loop
    from the vertices. The true angle of rotation between $X_a$ and
    $X_b$ is at $\frac{2\pi}{3}$. \textbf{(b-c)} Likelihood function
    and \naive{} (b) and BQ (c) model estimates. The sampled points
    $\mathbf{R}$ (red circles) are then used to estimate $S$ (black
    lines are the true $S$, red lines are the estimate).}
  \label{fig:shapes}
\end{figure}

We pick an initial direction of rotation which results in the higher
value of $S$. At each step, we compute $\mu_Z$ and $\tilde{V}$ to
estimate a distribution over the likelihood ratio $\ell$, i.e.
$p(\ell)\approx\frac{1}{p(X_b)}\ \mathcal{N}(Z\ \vert\ \mu_Z,
\sigma_z)$.  We choose $h_0$ when $p(\ell < 1)\geq 0.95$, and $h_1$
when $p(\ell > 1)\geq 0.95$. Until one of these conditions are met (or
the shape has been fully rotated), the model will continue to compute
rotations and update its estimate of $Z$.  The model may also change
direction based on the \textit{expected} posterior variance of $Z$
given some new sample $a$\footnote{This is similar to the full
  posterior variance calculation given in \cite{Osborne:2012tm},
  however we compute the variance given only the current mean estimate
  of $a$.}. The model will change directions if doing so would lower
this expected variance. Thus, it is able to actively change its
strategy, unlike the hill-climbing procedure.

\section{Results}

We evaluated each model's performance on 20 randomly generated shapes
which had between 3 and 6 vertices, inclusive, (e.g., $X_a$ in Fig.
\ref{fig:stimuli}). For each shape, we computed 18 ``same'' and 18
``different'' stimuli pairs, with $R$ spaced at $20^\prime$ increments
between 0 and 360, as in \cite{Shepard1971}. ``Same'' pairs were
created by rotating $X_a$ by $R$; the same was true for ``different'',
except that $X_a$ was also reflected across the $y$-axis. To gauge
performance, we looked at \textit{response error rates}: how accurate
was the model at choosing the correct hypothesis? This was defined as
the mean error ($\ME{}$), or fraction of times the model chose
incorrectly.  We additionally looked at \textit{rotations}: for those
``same'' pairs which the model judged correct, how correlated were the
model's rotations with the true angles of rotation?  We quantified
this using the Pearson's correlation coefficient $\rho$ for the true
rotation ($R$) versus the number of steps/rotations take by the model
($\vert \mathbf{R}\vert$). Fig. \ref{fig:rotations} shows true
rotations vs number of steps/rotations by the model for individual
stimuli (left) and average per true rotation (right). This analysis
can be qualitatively compared to the results of \cite{Shepard1971}
(Fig.  \ref{fig:mental-rotation}).

\paragraph{\Naive{}}

The \naive{} model's response error rate was $\NaiveME{}$, which is
better than chance (equivalent to guessing randomly,
i.e. $\ME{}=0.5$). The correlation between the \naive{} model's
average rotation and the true angle of rotation was $\Naivecorr{}$
(Fig. \ref{fig:rotations}, bottom left). For $R<\frac{\pi}{2}$, the
\naive{} model corresponds extremely well to the true angle of
rotation; this is because it needs to rotate less and is therefore
less likely to get stuck on local maxima. For $R>\frac{\pi}{2}$, we
see an increasing tendency to under-rotate due to getting stuck on
local maxima, as well as a tendency to over-rotate if the wrong
direction was initially chosen.

\paragraph{Bayesian Quadrature}

The BQ model was much more accurate in choosing the correct hypothesis
than the other model ($\BQME{}$). The number of rotations computed by
the BQ model were strongly correlated with the true rotations
($\BQcorr$), a result which is qualitatively similar to that exhibited
by humans (Fig. \ref{fig:mental-rotation}
vs. Fig. \ref{fig:rotations}). Because the BQ model has the capacity
to ``reset'', it could recover from rotating in the incorrect
direction (e.g., Fig. \ref{fig:bq}) and thus did not over-rotate as
frequently.  It also under-rotated less frequently: because the model
continues to rotate until it is confident in its estimate of $Z$, it
does not get stuck as easily on local optima.

\begin{figure}[t]
  \centering
  \includegraphics[width=\textwidth]{../figures/model_rotations.pdf}
  \caption{\textbf{Model rotations.} Left: each subplot shows the
    correspondence between the true angle of rotation ($R$) for
    ``same'' pairs and the amount of rotation performed by the
    model. Right: each subplot shows the models' mean rotations over
    stimuli pairs that were judged correctly. Black dots correspond to
    ``same'' pairs, and blue lines correspond to ``different''
    pairs. Error bars/shaded regions indicate one standard deviation,
    and the dotted lines indicate the least-squares fit to the
    ``same'' pairs.}
  \label{fig:rotations}
\end{figure}

\section{Conclusion}

In this paper, we asked: how do people use mental simulation?  We
investigated the specific case of mental rotation, using rational
analysis to characterize optimal strategies. We found that an
adaptive, BQ model performs best at the classic rotation task
\cite{Shepard1971}, as it is able to actively monitor the confidence
of its estimate and intelligently choose the direction of rotation. A
simpler model based on heuristics performed much worse: it did not
maintain the linear relationship with the true angle of rotation due
to lack of robustness against local maxima or incorrect rotation
direction.  More broadly, the BQ model provides answers to puzzling
questions surrounding the incremental nature of mental rotation: which
way should the object be rotated, and for how long? This model
formalizes these answers in a way that is qualitatively consistent
with human behavior, both in response time linearity
\cite{Shepard1971} and variability \cite{Just1976}. These results
further indicate that this may be another instance in which people do
appropriately use available computational resources to solve the task
at hand.  Rational, adaptive approaches such as this may therefore be
successful when applied to more complex cognitive processes above and
beyond mental rotation.

% \renewcommand\bibsection{\subsubsection*{\refname}}
\renewcommand\refname{\normalsize{References}}
\bibliographystyle{ieeetr}
{\small \bibliography{references}}

\end{document}