### 2014-04-30 man/cogsci-2014/cogsci-2014.tex
More tweaks, to fit on 6 pages

 
% Annual Cognitive Science Conference
% Sample LaTeX Paper -- Proceedings Format
%

% Original : Ashwin Ram (ashwin@cc.gatech.edu)       04/01/1994
% Modified : Johanna Moore (jmoore@cs.pitt.edu)      03/17/1995
% Modified : David Noelle (noelle@ucsd.edu)          03/15/1996
% Modified : Pat Langley (langley@cs.stanford.edu)   01/26/1997
% Latex2e corrections by Ramin Charles Nakisa        01/28/1997 
% Modified : Tina Eliassi-Rad (eliassi@cs.wisc.edu)  01/31/1998
% Modified : Trisha Yannuzzi (trisha@ircs.upenn.edu) 12/28/1999 (in process)
% Modified : Mary Ellen Foster (M.E.Foster@ed.ac.uk) 12/11/2000
% Modified : Ken Forbus                              01/23/2004
% Modified : Eli M. Silk (esilk@pitt.edu)            05/24/2005
% Modified: Niels Taatgen (taatgen@cmu.edu) 10/24/2006

%% Change ``a4paper'' in the following line to ``letterpaper'' if you are
%% producing a letter-format document.

\documentclass[10pt,letterpaper]{article}

%% For better looking math (but it takes up more space):
% \AtBeginDocument{\RequirePackage{lmodern, times}}

\usepackage{cogsci}
\usepackage{pslatex}
\usepackage{apacite}
\usepackage{graphicx}
\usepackage{color}
\usepackage{amsmath, amsthm, amssymb}
\usepackage{subcaption}
\usepackage{caption}

\newcommand{\TODO}[1]{\textcolor{red}{[TODO: #1]}}
\newcommand{\Xa}[0]{\mathbf{X}_a}
\newcommand{\Xb}[0]{\mathbf{X}_b}
\newcommand{\Xt}[0]{\mathbf{X}_t}
\newcommand{\R}[0]{\mathbf{R}_\theta}
\newcommand{\F}[0]{\mathbf{F}}
\newcommand{\M}[0]{\mathbf{M}}
\newcommand{\I}[0]{\mathbb{I}}
\newcommand{\hi}[0]{h=0}
\newcommand{\hf}[0]{h=1}
\newcommand{\dif}[0]{\,\mathrm{d}}

\newcommand{\Oc}[0]{Oracle}
\newcommand{\Th}[0]{Threshold}
\newcommand{\Hc}[0]{HC}
\newcommand{\Bq}[0]{BQ}

\newenvironment{pitemize}{
\begin{itemize}
  \setlength{\itemsep}{1pt}
  \setlength{\parskip}{0pt}
  \setlength{\parsep}{0pt}
}{\end{itemize}}

% space below "Figure 1: ...", but only for inline figures
%\addtolength{\textfloatsep}{-0.5cm}

\addtolength{\abovecaptionskip}{-0.25cm}
\addtolength{\belowcaptionskip}{-0.5cm}

\input{../../results/D/overall_response_time.tex}
\input{../../results/D/overall_accuracy.tex}
\input{../../results/D/num_chance.tex}
\input{../../results/D/trial_accuracy_corrs.tex}
\input{../../results/D/trial_time_corrs.tex}
\input{../../results/D/response_time_corrs.tex}
\input{../../results/D/accuracy_corrs.tex}
\input{../../results/D/human_corrs.tex}
\input{../../results/D/theta_time_corrs.tex}
\input{../../results/D/theta_accuracy_corrs.tex}

\title{What to simulate? Inferring the right direction for mental
  rotation}

\author{{\large \bf Jessica B. Hamrick (jhamrick@berkeley.edu)} \\
  {\large \bf Thomas L. Griffiths (tom\_griffiths@berkeley.edu)} \\
  Department of Psychology, University of California, Berkeley, CA
  94720 USA}

\begin{document}

\maketitle

\begin{abstract}
  When people use mental imagery, how do they decide \textit{which}
  images to generate? To answer this question, we explored how mental
  simulation should be used in the classic psychological task of
  determining if two images depict the same object in different
  orientations \cite{Shepard1971}. Through a rational analysis of
  mental rotation, we formalized four models and compared them to
  human performance. We found that three models based on previous
  hypotheses in the literature were unable to account for several
  aspects of human behavior. The fourth is based on the idea
  \textit{active sampling} \cite<e.g.,>{Gureckis:2012gu}, which is a
  strategy of choosing actions that will provide the most
  informative. This last model provides a plausible account of how
  people use mental rotation, where the other models do not. Based on
  these results, we suggest that the question of ``what to simulate?''
  is more difficult than has previously been assumed, and that an
  active learning approach holds promise for uncovering the answer.

  \textbf{Keywords:} mental rotation, computational modeling
\end{abstract}

\section{Introduction}

One of the most astonishing cognitive feats is our ability to
envision, manipulate, and plan with objects---all without actually
perceiving them. This \textit{mental simulation} has been widely
studied, including an intense debate about the underlying
representation of mental images \cite<e.g.,>{Kosslyn:2009tj,
  Pylyshyn:2002vk}. But this debate hasn't addressed one of the most
fundamental questions about mental simulation: how people decide
\textit{what} to simulate.

Mental rotation provides a simple example of the decision problem
posed by simulation.  In the classic experiment by
\citeA{Shepard1971}, participants viewed images of three-dimensional
objects and had to determine whether the images depicted the same
object (which differed by a rotation) or two separate objects (which
differed by a reflection and a rotation). They found that people's
response times (RTs) had a strong linear correlation with the minimum
angle of rotation, a result which led to the conclusion that people
solve this task by ``mentally rotating'' the objects until they are
congruent.  However, this explanation leaves several questions
unanswered. How do people know the axis around which to rotate the
objects? If the axis is known, how do people know which direction to
rotate the objects?  And finally, how do people know how long to
rotate?

In this paper, we explore these questions through rational analysis
\cite{Marr:1983to,anderson90,Shepard:1987tt} and compare four models
of mental rotation. We begin the paper by discussing the previous
literature on mental imagery. Next, we outline computational- and
algorithmic-level analyses of the problem of mental rotation.  We then
describe a behavioral experiment based on the classic mental rotation
studies \cite<e.g.,>{Cooper:1975wp}, and compare the results of our
experiment with each of the models. We conclude with a discussion of
the strengths and weaknesses of each model, and lay out directions for
future work.

\section{Modeling mental rotation}

Previous models of mental rotation have largely focused on the
representation of mental images, rather than how people decide
\textit{which} mental images to generate. \citeA{Kosslyn:1977tv}
proposed a model of the mental imagery buffer, but did not say
\textit{how} it should be used. Similarly, \citeA{Julstrom:1985va} and
\citeA{Glasgow:1992tj} were mostly concerned with modeling the
representational format underlying imagery. Although
\citeA{Anderson1978} emphasized the importance of considering both
representation and process, he dismissed the problem of determining
the direction of rotation as a ``technical difficulty''.

The only models (of which the authors are aware) that seriously
attempted to address the decision of \textit{what} to simulate are
those by \citeA{Funt:1983wn} and \citeA{Just:1985uu}. In both of these
models, the axis and direction of rotation are computed prior to
performing the rotation. One object is then rotated through the target
rotation, and is checked against the other object for
congruency. However, this approach assumes that the corresponding
points on the two objects can be easily identified, which is not
necessarily the case.  Indeed, the state-of-the-art in computer vision
suggests that there is more to this problem than checking for
congruency, particularly when the shapes are complex or not exactly
the same \cite<e.g.,>{Belongie:2002tj,Sebastian:2003vm}. Additionally,
recent research shows that when performing \textit{physical}
rotations, people do not rotate until congruency is reached; they may
even rotate \textit{away} from near perfect matches
\cite{Gardony:2013gn}.

If people are not computing the rotation beforehand, what might they
be doing? To answer this question, we perform a rational analysis of
the problem of mental rotation
\cite{Marr:1983to,anderson90,Shepard:1987tt}. At the computational
level, we can say that the \textit{problem} is to determine which
spatial transformations an object has undergone based on two images of
that object (which do not include information about point
correspondences). At the algorithmic level, we are constrained by the
notion that mental images must be transformed in an analog manner (or
in a way that is approximately analog), and that mental images are
time-consuming and effortful to generate. Thus, the \textit{goal} is
to make this determination while performing a minimum amount of
computation (i.e., as few rotations as possible).

The original ``congruency'' hypothesis \cite{Shepard1971} is a
rational solution to this problem, in the sense that the smallest
amount of computation coincides with rotating through the minimum
angle. However, it violates the constraint that we do not know the
points of correspondence between the images, which is what
necessitates the use of imagery. Noting that a rational solution need
not maintain a single trajectory of rotation, we explore an
alternative model, which---rather than computing the angle of
rotation---engages in an \textit{active sampling} strategy.

Active sampling is the idea people gather new information in a manner
that increases certainty about the problem space. An everyday example
of this can be observed in the game of ``20 questions'', in which one
person thinks of a concept, and another has to guess the concept in 20
questions or less. The first question is almost always ``person,
place, or thing?'', because the answer provides the most possible
information about the concept of interest. Active sampling has
gained support across several areas of cognitive science
\cite<e.g.,>{Gureckis:2012gu}, including other spatial domains
\cite{Juni:2011vo}. In the case of mental rotation, actively choosing
rotations may be the best way to gather evidence about the similarity
between the observed shapes when the angle of rotation is unknown.

\section{How should we rotate?}

In this section, we formalize our rational analysis and propose four
models of mental rotation: one based on existing models; two which are
extensions of the first but with relaxed assumptions; and one based on
the active sampling approach.

The task we are interested in modeling involves observing two images
and determining whether one image depicts the ``same'' object as the
other image (differing by a rotation), or a ``flipped'' version of the
object in the other image (differing by a reflection and then a
rotation).

\subsection{Computational-level analysis}

\begin{figure}[t]
  \begin{center}
    \includegraphics[width=0.48\textwidth]{../../figures/stimuli_and_similarity.pdf}
    \caption{\textbf{Example stimuli and similarities.} This figure
      shows a ``flipped'' stimulus pair with a rotation of
      $120^\circ$, and the corresponding similarity functions for each
      hypothesis. Arrows indicate where each shape lies on the curve.}
    \label{fig:shapes}
  \end{center}
\end{figure}

We denote the shapes as $X_a$ and $X_b$ and assume $X_b$ is generated
by a transformation of $X_a$, i.e. $X_b=f(X_a, \theta, h)$, where
$\theta$ is a rotation, $\hi$ is the hypothesis that the images depict
the same object, and $\hf$ is the hypothesis that the images depict
mirror-image objects. The posterior probability of each hypothesis
given the observed shapes is then:
$p(h\ \vert\ X_a, X_b) \propto \int p(X_b\ \vert\ X_a, \theta,
h)p(h)p(\theta)\dif\theta$,
where $p(X_b\ \vert\ X_a, \theta, h)$ is the probability that $X_b$
was generated from $X_a$. Because we want to determine which
hypothesis is more likely, the quantity of interest is a ratio
$\ell:=p(\hi\ \vert\ X_a, X_b) / p(\hf\ \vert\ X_a, X_b)$ which
(assuming that all rotations are equally likely) is equivalent to:
\begin{equation}
  \ell = \frac{\left(\int p(X_b\ \vert\ X_a, \theta, \hi)\dif\theta\right)\cdot{}p_0}{\left(\int p(X_b\ \vert\ X_a, \theta, \hf)\dif\theta\right)\cdot{}p_1},
  \label{eq:lh-ratio}
\end{equation}
where $p_0=p(h=0)$ and $p_1=p(h=1)$, for brevity. If $\ell > 1$, then
we accept the hypothesis that the images depict the same object
($\hi$); if $\ell < 1$, then we accept the hypothesis that the images
depict flipped objects ($\hf$).

\subsection{Algorithmic constraints}

\begin{figure*}[t]
  \begin{center}
    \includegraphics[width=\textwidth]{../../figures/D/response_time_accuracy.pdf}
    \caption{\textbf{Response time and accuracy comparison.} Top: RT
      of correct responses as a function of the minimum angle of
      rotation. Bottom: accuracy as a function of the minimum angle of
      rotation. All error bars are 95\% confidence intervals.}
    \label{fig:response-time-accuracy}
  \end{center}
\end{figure*}

We represent a shape of $N$ vertices with a $N\times 2$ coordinate
matrix $\mathbf{X}=[\mathbf{x}_1, \ldots{}, \mathbf{x}_N]$, and denote
the rotation and/or reflection transformation as $f(\mathbf{X}, h,
\theta):=\mathbf{X}\F_h^T\R^T$, where $\R$ is a rotation matrix, and
$\F_h$ is either the identity matrix $\I$ (when $\hi$) or a reflection
matrix across the $y$-axis (when $\hf$).

We define $p(\Xb\ \vert\ \Xa, \theta, h)$ to be the similarity between
$\Xb$ and a transformation of $\Xa$: $p(\Xb\ \vert\ \Xa, \theta, h):=
S(\Xb, f(\Xa, h, \theta))$.  We do not know which vertices of $\Xb$
correspond to which vertices of $\Xa$, so the similarity $S$ must
marginalize over the set of possible mappings. For brevity, let
$\mathbf{X}_m=\M\cdot{}f(\Xa, h, \theta)$ where $\M$ is a permutation
matrix. Then:
\begin{equation}
  S(\Xb, f(\Xa, h, \theta)):=\frac{1}{2N} \sum_{\M} \prod_{n=1}^N \mathcal{N}(\mathbf{x}_{bn}\ \vert \ \mathbf{x}_{mn}, \I\sigma_S^2),
  \label{eq:similarity}
\end{equation}
where $2N$ is the total number of possible mappings,\footnote{It is
  $2N$ and not $N^2$ because, in polar coordinates, vertices are
  always connected to their two nearest neighbors in the $\theta$
  dimension.}and $\sigma_S^2=0.15$ is the variance of the
similarity. Example similarity curves are shown in
Figure~\ref{fig:shapes}.

We assume that the observed shapes must be transformed by a small
amount at a time, and each transformation takes a non-negligible
amount of time. If the current mental image is $\Xt$, then:
\begin{equation}
  \mathbf{X}_{t+1} = \left\{ \begin{array}{ll}
      f(\Xt, 0, \epsilon) &\mbox{ rotate by $\epsilon$ radians,} \\
      f(\Xt, 1, 0) &\mbox{ flip,} \\
      f(\Xa, 0, 0) &\mbox{ reset to $0^\circ$, or} \\
      f(\Xa, 1, 0) &\mbox{ reset and flip,} \\
    \end{array} \right.
  \label{eq:actions}
\end{equation}
where $\epsilon\sim \left|\mathcal{N}(0, \sigma_\epsilon^2)\right|$
and $\sigma_\epsilon^2$ is the variance of the step size.

To summarize, we approximate the likelihood term of
Equation~\ref{eq:lh-ratio} using the similarity function defined in
Equation~\ref{eq:similarity}. Because we assume mental rotations are
performed sequentially, this similarity can only be computed for the
actions listed in Equation~\ref{eq:actions}.

\subsection{Specific models of mental rotation}

In order to approximate Equation~\ref{eq:lh-ratio} using samples of
the similarity function, we must decide \textit{which} places to
sample and \textit{when} stop sampling. The models below differ in how
they make these decisions.

\subsubsection{Oracle model}

One hypothesis is that people compute the direction and extent of
rotation beforehand using \textit{a priori} knowledge of the
correspondence between points in the images
\cite{Funt:1983wn,Just:1985uu}.  To reflect this hypothesis, we
created an ``oracle'' model which is told which points on each shape
correspond. From that correspondence, it computes the correct rotation
and rotates through it.

To determine the correct rotation, we solve for the rotation matrix by
computing $(\Xa \F_h^T)_\mathrm{left}^{-1}\cdot{}\Xb$, where
$(\Xa\F_h^T)_\mathrm{left}^{-1}$ is the left inverse of
$\Xa\F_h^T$. We then check each $h$ to see if the computation produces
a valid rotation matrix; the $h$ that does is the correct
hypothesis. This gives us the true value of $\theta$, so
Equation~\ref{eq:lh-ratio} becomes a generalized likelihood ratio
test, where $\theta$ is set to the MLE value, rather than being
marginalized:
\begin{equation}
  \ell = \frac{\max_\theta p(\Xb\ \vert\ \Xa, \theta, \hi)\cdot{}p_0}{\max_\theta p(\Xb\ \vert\ \Xa, \theta, \hf)\cdot{}p_1}.
  \label{eq:mle-lh-ratio}
\end{equation}

If we give equal weight to the two hypotheses, then the priors cancel
out; if we weigh one hypothesis more heavily, then our decision will
be biased towards that hypothesis. However, unless the likelihood
ratio is already very close to 1, small biases in the prior will not
make much of a difference.

\subsubsection{Threshold model}

A model which does not know point correspondences could use the
following algorithm: (1) pick a random direction; (2) take a single
step; (3) if that step decreased similarity, then begin rotating in
the reverse direction, otherwise continue rotating in the original
direction; (4) continue rotating in the chosen direction until a
``match'' is found (defined as finding a value of $S$ that exceeds a
threshold); and (5) if no match was found, flip, and start over from
step one.  We only allow for the ``flip'' action after no match has
been found, because there is no particularly principled way for the
\Th{} model to choose when to flip. We assume that the locations where
$S$ is greater than the threshold correspond to the true $\theta$ (or
points near the true $\theta$). So, as with the \Oc{} model, we use
Equation~\ref{eq:mle-lh-ratio}.

\subsubsection{Hill Climbing model}

\begin{figure*}[t]
  \begin{center}
    \includegraphics[width=\textwidth]{../../figures/D/response_time_scatters.pdf}
    \caption{\textbf{Model vs. human RTs.} Each subplot shows the
      z-scored model RTs ($x$-axis) vs. the z-scored human RTs
      ($y$-axis). Pearson correlations are shown beneath each
      subplot. The dotted lines are $x=y$.}
    \label{fig:human-model-scatters}
  \end{center}
\end{figure*}

In the current formulation of the problem, choosing the threshold is
straightforward because we know both the exact geometry of the shapes
and that a linear transformation exists which will align
them. However, this choice is not always clear \textit{a priori}, as
the global optimum depends on many factors (e.g., shape complexity,
dimensionality, perceptual uncertainty, and whether the shapes are
identical).  One way to deal with the problem of choosing a threshold
would be use a global optimization strategy; however, this would not
result in the linear RT found by \citeA{Shepard1971}. A second
alternative is to perform a Hill Climbing (\Hc{}) search; i.e., rotate
in the direction that increases similarity until no further
improvement can be found. In contrast with the \Th{} model, this
results in arriving in a \textit{local} maximum (which may or may not
be the global maximum). Thus, as with the \Oc{} and \Th{} models, we
use Equation~\ref{eq:mle-lh-ratio}. We only allow for the ``flip''
action after a local maximum has been reached, because like the \Th{}
model, there is otherwise no principled way for the \Hc{} model to
choose when to flip.

\subsubsection{Bayesian Quadrature model}

While the previous few models all focused on \textit{searching} for
the global maximum, we need only \textit{approximate}
Equation~\ref{eq:lh-ratio}. We hypothesize a model based on the idea
of \textit{active sampling} \cite<e.g.,>{Gureckis:2012gu}: instead of
searching for a maximum, we maintain a probability distribution over
our \textit{estimate} of Equation~\ref{eq:lh-ratio}, and then sample
actions which are expected to improve that estimate.  This strategy
has the benefits that it does not make assumptions about the scale of
the similarity function; and, by choosing to sample places which are
informative, this method implicitly minimizes the amount of rotation.

We denote $Z_h$ as our estimate of the likelihood for hypothesis $h$,
and write its distribution as:
$p(Z_h) = \int \left[\int S(\Xb, f(\Xa, \theta,
  h))p(\theta)\dif\theta\right] p(S)\dif S$,
where $S$ is the similarity function, and $p(S)$ is a prior over
similarity functions.  This method of estimating an integral is known
in the machine-learning literature as \textit{Bayesian Quadrature}
\cite{Diaconis:1988uo,Osborne:2012tm}, or BQ.  Denoting
$S_h=S(\Xb, f(\Xa, \theta, h))$, we first place a \textit{Gaussian
  Process} \cite{Rasmussen:2006vz}, or GP, prior on the log of $S_h$
in order to enforce positivity after it is exponentiated, i.e.
$\mathbb{E}[Z_h] \approx \int \exp(\mu_h(\theta))p(\theta)\dif\theta$,
where $\mu_h:=\mu(\log S_h)$ is the mean of the log-GP
\cite{Osborne:2012tm}.  To approximate this integral, we fit a second
GP over points sampled from the log-GP, which we denote as
$\bar{S}_h:=\exp(\mu_h)$. Then, from \citeA{Duvenaud:2013td}, we have
$\mathbb{E}[Z_h] \approx \int \bar{\mu}_h(\theta)p(\theta)\dif\theta$
and
$\mathbb{V}(Z_h) \approx \iint \mathrm{Cov}_h(\theta,
\theta^\prime)\bar{\mu}_h(\theta)\bar{\mu}_h(\theta^\prime)p(\theta)p(\theta^\prime)\dif\theta\dif\theta^\prime$,
where $\bar{\mu}_h:=\mu(\bar{S}_h)$ is the mean of the second GP, and
$\mathrm{Cov}_h:=\mathrm{Cov}(\log S_h)$ is the covariance of the
log-GP.

Assuming independence, we can now write
$p(Z_h)\approx\mathcal{N}(Z_h\ \vert\ \mathbb{E}[Z_h],
\mathbb{V}(Z_h))$,
which gives us a distribution over the likelihood ratio in
Equation~\ref{eq:lh-ratio}:
$p(\ell)\approx\mathcal{N}(Z_0\ \vert\ \mathbb{E}[Z_0],
\mathbb{V}(Z_0))\cdot{}p_0/\mathcal{N}(Z_1\ \vert\ \mathbb{E}[Z_1],
\mathbb{V}(Z_1))\cdot{}p_1$.
This distribution cannot easily be calculated, but we are only
interested in whether $Z_0>Z_1$ or $Z_1>Z_0$. So, we use
$Z_D=p_0\cdot{}Z_0-p_1\cdot{}Z_1$ and compute
$p(Z_D)\propto\mathcal{N}(p_0\cdot{}\mathbb{E}[Z_0] -
p_1\cdot{}\mathbb{E}[Z_1], p_0^2\cdot{}\mathbb{V}(Z_0) +
p_1^2\cdot{}\mathbb{V}(Z_1))$.
We then sample new observations until we are at least 95\% confident
that $Z_D\neq 0$. In other words, when $p(Z_D<0)<0.025$, we accept
$\hi$, and when $p(Z_D<0)>0.975$, we accept $\hf$. Because we compare
the hypotheses in order to determine when to stop sampling, biasing
the prior should result in requiring less evidence for one hypothesis
before stopping, and more evidence for the other hypothesis.

To choose where to sample, we compute the expected variance of $Z_h$
given a new observation at $\theta_a$. From \citeA{Osborne:2012tm}, we
compute
$\mathbb{E}[\mathbb{V}(Z_h|\theta_a)]=\mathbb{V}(Z_h) +
\mathbb{E}[Z_h] - \int \mathbb{E}[Z_h|\theta_{a}]^2
\mathcal{N}(\mu_h(\theta_a), \mathrm{Cov}_h(\theta_a,
\theta_a))\dif\log S_h(\theta_a)$
for each of the actions in Eq.~\ref{eq:actions}; we pick the one with
the lowest value.

\section{Methods}

To evaluate the models described previously, we ran a behavioral
experiment based on classic mental rotation studies
\cite<e.g.>{Shepard1971, Cooper:1975wp}.

\subsubsection{Stimuli}

We randomly generated 20 shapes of five or six vertices (e.g.,
Figure~\ref{fig:shapes}). For each shape, we computed 20 ``same'' and
20 ``flipped'' stimuli pairs, with 18 rotations ($\theta$) spaced at
$20^\circ$ increments between $0^\circ$ and $360^\circ$ (with
$0^\circ$ and $180^\circ$ repeated twice, in order to gather an equal
number of responses for each angle between $0^\circ$ and
$180^\circ$). ``Same'' pairs were created by rotating $\Xa$ by
$\theta$; ``flipped'' pairs were first reflected $\Xa$ across the
$y$-axis, then rotated by $\theta$.

We generated five additional shapes to be used in a practice block of
10 trials. Across these trials, there was one ``flipped'' and one
``same'' repetition of each shape and each angle ($60^\circ$,
$120^\circ$, $180^\circ$, $240^\circ$, or $300^\circ$) such that no
shape was presented at the same angle twice. We also generated a sixth
shape to include with the instructions.  This shape had both a
``flipped'' and ``same'' version, each rotated to $320^\circ$.

\subsubsection{Participants and Design}

We recruited 247 participants on Amazon's Mechanical Turk using the
psiTurk experiment framework \cite{Mcdonnell12}. Each participant was
paid \$1.00 for 15 minutes of work, consisting of one block of 10
practice trials followed by two blocks of 100 randomly ordered
experiment trials.

All participants saw the same 10 practice trials as described
above. There were 720 unique experimental stimuli (20 shapes $\times$
18 angles $\times$ 2 reflections), though because stimuli with
rotations of $0^\circ$ or $180^\circ$ were repeated twice, there were
800 total experimental stimuli. These stimuli were split across eight
conditions in the following manner: first, stimuli were split into
four blocks of 200 trials. Within each block, each shape was repeated
ten times and each rotation was repeated ten times (five ``same'',
five ``flipped''), such that across all blocks, each stimulus appeared
once. Each block was then split in half, and participants completed
two half-blocks.

\subsubsection{Procedure}

\begin{figure*}[t]
  \begin{center}
    \includegraphics[width=\textwidth]{../../figures/D/response_time_histograms.pdf}
    \caption{\textbf{Response time histograms.} Each subplot shows the
      distribution of RTs on correct trials for people and the
      models.}
    \label{fig:histograms}
  \end{center}
\end{figure*}

Participants were given the following instructions while being shown
an example ``same'' pair and an example ``flipped'' pair: \textit{``On
  each trial, you will see two images. Sometimes, they show the
  \textbf{same} object. Other times, the images show \textbf{flipped}
  objects. The task is to determine whether the two images show the
  \textbf{same} object or \textbf{flipped} objects.''}

On each trial, participants were instructed to press the `b' key to
begin and to focus on the fixation cross that appeared for 750ms
afterwards. The two images were then presented side-by-side, each at
300px $\times$ 300px, and participants could press `s' to indicate
they thought the images depicted the ``same'' object, or `d' to
indicate they thought the images depicted ``flipped'' objects.  While
there was no limit on RT, we urged participants to answer as quickly
as possible while maintaining at least 85\% accuracy in the
experimental blocks.

\section{Results}

Of the 247 participants, 200 (81\%) were included in our analyses. Of
the other 47, we excluded 10 (4\%) because of an experimental error, 6
(2.4\%) because they had already completed a related experiment, and
31 (12.6\%) because they failed a comprehension check, which was
defined as correctly answering at least 85\% of stimuli with a
rotation of either $0^\circ$, $20^\circ$, or $340^\circ$.  We also
excluded 82 trials for which the RT was either less than 100ms or
greater than 20s.

For each model, we ran 50 samples for each of the 800 experimental
stimuli. The step size parameter ($\sigma_\epsilon$) was fit to human
RTs for each of the models, resulting in $\sigma_\epsilon=0.6$ for the
\Th{} and \Bq{} models and $\sigma_\epsilon=0.1$ for the \Oc{} and
\Hc{} models. We also ran the models under two different priors,
$p(h=0)=0.5$ (the ``equal'' prior) and $p(h=0)=0.55$ (the ``unequal''
prior). As expected, this only had a major effect on the stopping
criteria for the \Bq{} model.

\subsubsection{General analysis}

For analyses of RT, confidence intervals around harmonic means of
correct responses were computed using a bootstrap analysis of 10000
bootstrap samples (sampled with replacement).  We also used a
bootstrap analysis of 10000 bootstrap samples to compute the
confidence intervals around both Spearman ($\rho$) and Pearson ($r$)
correlations.  Unless otherwise specified, all correlations were
computed over 720 stimuli. For analyses of accuracy, confidence
intervals were computed from a binomial proportion with a Jeffrey's
beta prior.  To test if judgments were above chance on a particular
stimulus, we used the same binomial proportion and tested whether
$p\left(p(\textrm{correct})\leq 0.5\right)\leq \frac{0.05}{720}$,
where $\frac{1}{720}$ is a Bonferroni correction for multiple
comparisons.

\subsubsection{Human}

The average RT across all correctly-judged stimuli was \ExpTime{} ms;
the full histogram of RTs can be seen in
Figure~\ref{fig:histograms}. The minimum angle of rotation was
significantly rank-order (Spearman) correlated with average
per-stimulus RTs, both for ``flipped'' (\ExpThetaTimeCorrFlipped{})
and ``same'' pairs (\ExpThetaTimeCorrSame{}). While this replicates
the general result of previous experiments
\cite<e.g.,>{Shepard1971,Cooper:1975wp}, our results are not as linear
(Figure~\ref{fig:response-time-accuracy}).\\
\indent The average accuracy across all stimuli was \ExpAccuracy{},
though there were \ExpNumChance{} stimuli (out of 720) for which
people were not above chance.  The minimum angle was also correlated
with participants' average per-stimulus accuracy, though much more so
for ``same'' pairs (\ExpThetaAccuracyCorrSame{}) than ``flipped''
pairs (\ExpThetaAccuracyCorrFlipped{}). This is the same
result found both by \citeA{Cooper:1975wp} and \citeA{Gardony:2013gn}.\\
\indent There was a significant effect of trial number both on RT
(\ExpTrialTimeCorr{}) and on accuracy (\ExpTrialAccuracyCorr{}),
though the effect on accuracy was not significant during the second
half of the experiment (\ExpaTrialAccuracyCorr{} for the first half
vs. \ExpbTrialAccuracyCorr{} for the second half). These effects may
have contributed to the not-quite-linearity of the human RTs; future
work should collect more data per participant.

\subsubsection{\Oc{} model}

The number of actions taken by the \Oc{} model was perfectly
correlated with the minimum angle of rotation
(Figure~\ref{fig:response-time-accuracy}). The \Oc{} model was the
best fit to human RTs, with a correlation of \OcTimeCorr{}
(Figure~\ref{fig:human-model-scatters}), although the distribution of
response times did not match that of people
(Figure~\ref{fig:histograms}). Moreover, the \Oc{} model was 100\%
accurate, and therefore could not explain the effect of rotation on
people's accuracy.

\subsubsection{\Th{} model}

There was an overall monotonic relationship between the minimum angle
of rotation and the number of actions taken by the \Th{} model
(Figure~\ref{fig:response-time-accuracy}), though this relationship
did not hold for \textit{individual} shapes (e.g.,
Figure~\ref{fig:response-time-stimulus}).  The \Th{} model was able to
explain a moderate amount of the variance in human RTs, with a
correlation of \ThTimeCorr{}
(Figure~\ref{fig:human-model-scatters}). Like the \Oc{} model, the
overall distribution of its RTs did not match that of people
(Figure~\ref{fig:histograms}).  The \Th{} model had 100\% accuracy,
and thus did not exhibit a relationship between minimum angle and
accuracy.  As noted, we fit $\sigma_\epsilon=0.6$ for the \Th{}
model. This had the interesting effect of causing the \Th{} model to
\textit{over}rotate, because the step size was large enough that it
sometimes missed the global maximum, and had to do another full
rotation to find it.

\subsubsection{\Hc{} model}

The \Hc{} was the only model for which there was no monotonic
relationship between rotation and RT
(Figure~\ref{fig:response-time-accuracy}). Moreover, the \Hc{} model
was barely above chance (\HcAccuracy{}) and there were \HcNumChance{}
stimuli for which it was not above chance. The \Hc{} model was not a
good predictor of human RTs (\HcTimeCorr{}), as shown in
Figure~\ref{fig:human-model-scatters}. It was a moderate predictor of
human accuracy (\HcAccuracyCorr{}).

\subsubsection{\Bq{} model}

Like the \Oc{} and \Th{} models, there was an overall monotonic
relationship between rotation and the number of steps taken by the
\Bq{} model (Figure~\ref{fig:response-time-accuracy}). Unlike the
\Th{} model, this relationship existed for individual shapes as well
(e.g., Figure~\ref{fig:response-time-stimulus}).  The \Bq{} model
explained variance in human RTs about as well as the \Th{} model
(Figure~\ref{fig:human-model-scatters}), with a correlation of
\BqTimeCorr{} for the equal prior and \BqpTimeCorr{} for the unequal
prior, and the RT distribution from the \Bq{} model had the same
overall shape as that of people (Figure~\ref{fig:histograms}).\\
\indent The \Bq{} model was quite accurate overall (equal prior:
(\BqAccuracy{}; unequal prior: \BqpAccuracy{}). With the equal prior,
there were \BqNumChance{} stimuli for which it was not above chance;
with the unequal prior, there were \BqpNumChance{}. The correlation
with people's accuracy was \BqAccuracyCorr{} (equal prior) and
\BqpAccuracyCorr{} (unequal prior).\\
\indent Because the \Bq{} model relies on Equation~\ref{eq:lh-ratio}
for its stopping criteria (as opposed to just finding a maximum), the
prior $p(h)$ had an observable effect
(Figure~\ref{fig:response-time-accuracy}). As expected, with just a
small bias of $p(h=0)=0.55$, there was a clear separation in RTs for
``same'' versus ``flipped'' stimuli: because of this bias, the model
needed less evidence before accepting $h=0$ (thus taking less
time). This separation is similar to the trend also observed in human
RTs. The prior also had an effect on accuracy (though this did not
reflect human behavior): the bias towards $h=0$ meant that the model
was more likely to judge a pair as ``same'', thus, accuracy increased
for ``same'' pairs, but decreased for ``flipped'' pairs.

\section{Discussion}

\begin{figure}[t]
  \begin{center}
    \includegraphics[width=0.48\textwidth]{../../figures/D/response_time_stimulus.pdf}
    \caption{\textbf{Typical RT curves for a single object.}  These
      plots correspond to the object shown in
      Figure~\ref{fig:shapes}. Left: human curves are either linear
      (as with the ``same'' pairs), or linear and then flat (as with
      the ``flipped'' pairs). Middle: the \Th{} model does not have a
      monotonic relationship with rotation. Right: the \Bq{} model is
      roughly linear.}
    \label{fig:response-time-stimulus}
  \end{center}
\end{figure}

We set out to answer the question of how people decide \textit{what}
to simulate when using mental imagery. Focusing on the specific case
of determining the direction and extent of mental rotation, we
formalized four models and compared their performance with the results
of a behavioral experiment.

The \Oc{} and \Th{} models were the best predictors of human
RTs. However, both are somewhat unsatisfying explanations because they
rely on \textit{a priori} knowledge that people are unlikely to
have. Moreover, they offer no explanation of several aspects of human
behavior. First, their overall RT distributions look nothing like
people's (Figure~\ref{fig:histograms}).  Second, they both are 100\%
accurate, and so cannot explain the systematic relationship between
rotation and human accuracy
(Figure~\ref{fig:response-time-accuracy}). Third, neither model can
explain the difference in people's behavior on ``same'' and
``flipped'' stimuli.

In contrast, the \Bq{} model was nearly as good as the \Th{} model,
yet it makes no assumptions about people's \textit{a priori}
knowledge. Furthermore, the \Bq{} model matches people's behavior
better than the \Oc{} or \Th{} models in several ways. Its overall RT
histogram has the same general shape as people's
(Figure~\ref{fig:histograms}).  Moreover, a closer look shows that the
\Bq{} model maintains the monotonic relationship between angle and RT
even on individual stimuli, while the \Th{} model does not
(Figure~\ref{fig:response-time-stimulus}). Finally, the \Bq{} model's
adaptive stopping rule is sensitive to the prior, and thus provides a
possible explanation for why people are slower to respond on
``flipped'' stimulus pairs.

Thus, we suggest that the \Bq{} model offers the most promising
explanation of people's behavior on the mental rotation task to
date. While it is not a perfect account, there are several ways in
which it could be improved. For example, while we used holistic
rotations in this paper, there is evidence that people compare
individual features of shapes
\cite{Just1976,Yuille:1982tx}. Additionally, a different active
sampling approach could maintain a distribution over the location and
value of the global maximum, rather than over the integral. We intend
to explore these possibilities in future work, building upon the
foundation established in this paper and working towards a better
understanding of \textit{what} people choose to simulate.

%% TODO: Uncomment this for final version:

\subsubsection{Acknowledgments} {\small This research was supported by ONR
  MURI grant number N00014-13-1-0341, and a Berkeley Fellowship
  awarded to JBH. }

\bibliographystyle{apacite}
\renewcommand{\bibliographytypesize}{\small}
\setlength{\bibleftmargin}{.125in}
\setlength{\bibindent}{-\bibleftmargin}
\bibliography{references}

\end{document}